---
title: "Gender Equity in Louisville"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      dev.args=list(bg="transparent"), fig.width=16,fig.height=12)



library(glptools)
glp_load_packages(TRUE)
library(arrow)
library(viridis)
library(forcats)
library(hrbrthemes)
library(glpdata)


showtext_auto()
font_add("Museo Sans", "MuseoSans_300.otf")

```

```{r, include=FALSE}
#Overlayed histogram
func_plt_hist_overlay <- function(data, var) {

  data$group <- data[[var]]

  possible_colors <- c("#0E4A99", "#F58021", "#00A9B7", "#800055", "#356E39", "#CFB94C", "#7E9C80")
  number_I_need = length(unique(data$group))
  these_colors <- possible_colors[1:number_I_need]

  p <- data %>% ggplot( aes(x=HHINCOME, fill=group, color = group, weight = HHWT)) +
    geom_histogram(alpha=0.5, position = 'identity', binwidth = 10000) +
    scale_fill_manual(values = these_colors) +
    scale_color_manual(values = these_colors) +

    labs(fill="") +
    xlab("Household Income") +
    ylab("Percentage") 
    
  
  return (p)
  
}

glp_graph_theme <- theme_bw(
    base_size = 8,
    base_family = "Museo Sans")

glp_graph_theme <- glp_graph_theme +
  theme(
    legend.title     = element_blank(),
    legend.position  = "top",
    legend.margin    = margin(t = 10, unit = "pt"),
    legend.spacing.x = unit(30, "pt"),
    legend.text      = element_text(size = 40,
                                    margin = margin(b = 15, t = 15, unit = "pt")),

    axis.text    = element_text(size = 50),
    axis.title   = element_text(size = 60),
    axis.title.x = element_text(margin = margin(t = 15, unit = "pt")),
    axis.title.y = element_text(margin = margin(r = 15, unit = "pt")),

    plot.title = element_text(size = 90,
                              hjust = .5,
                              margin = margin(b = 25, unit = "pt")),

    plot.caption = element_text(size = 40,
                                lineheight = 0.5)) +
    theme(
      panel.background = element_rect(fill = "transparent", color = NA), # bg of the panel
      plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
      legend.background = element_rect(fill = "transparent", color = "transparent"), # get rid of legend bg
      legend.box.background = element_rect(fill = "transparent", color = "transparent"), # get rid of legend panel bg
      legend.key = element_rect(fill = "transparent",colour = NA)
    )
  
      # theme(plot.subtitle = element_text(hjust = 0.5, size = 50)) +
      # labs(subtitle = subtitle_text)

```

```{r prep_data, eval=FALSE, include=FALSE}
#Skip
# Harrison
#census_data <- read_feather("../../glpdata/data-raw/microdata/acs_micro_repwts.feather")

#Josh
#census_data <- read_feather("/Users/joshmarkwell/Desktop/Professional/GLP/projects/gender_equity/raw_data/acs_micro_repwts.feather")


census_data %<>%
  filter(FIPS == "21111")

write_feather(census_data, "raw_data/census_data.feather")


save(population_tract, file = "raw_data/tract_population.RData")

```




```{r read in new data, eval=FALSE, include=FALSE}
#Skip - only use if changing data to include

#acs_micro_8_11_22 <- read_feather("raw_data/acs_micro_FIPS_repwts_8_11_22.feather") 

census_data %<>%
  filter(FIPS == "21111" | FIPS == '37081')


write_feather(census_data, "raw_data/census_data_8_11_22.feather")

```


```{r explore new data, include=FALSE}
#Start
#File for use below
census_microdata081122_person <- read_feather("raw_data/census_data_8_11_22.feather") 

```



```{r clean data, include= FALSE}
# To classify the housing data into homeowners, renters, cost-bu
#also create income brackets, earner types
if ("census_microdata081122_processed.feather" %in% list.files("raw_data")) {
  census_microdata081122 <- arrow::read_feather("raw_data/census_microdata081122_processed.feather")
  census_microdata081122_person <- arrow::read_feather("raw_data/census_microdata081122_person_processed.feather")
} else {
  
  census_microdata081122_person %<>%
    group_by(year, SERIAL) %>%
    mutate(hh_members = n()) %>%
    ungroup() %>%
    #filter(PERNUM == 1) %>%
    mutate(
      OWNCOST  = replace(OWNCOST, OWNCOST == 99999, NA),
      HHINCOME = replace(HHINCOME, HHINCOME == 99999, NA),
      OWNERSHP = replace(OWNERSHP, OWNERSHP == 0, NA),
      RENTGRS  = replace(RENTGRS, RENTGRS == 0 & OWNERSHP == 1, NA),
  
      homeownership = if_else(OWNERSHP == 1, T, F),
  
      hcost = if_else(homeownership == 1, OWNCOST, RENTGRS),
      cost_burden = if_else(hcost * 12 / HHINCOME > 0.3, T, F),
      severe_cost_burden = if_else(hcost * 12 / HHINCOME > 0.5, 1, 0),
  
      hh_type = case_when(
        homeownership  & !cost_burden ~ "noncb_homeowner",
        homeownership  & cost_burden  ~ "cb_homeowner",
        !homeownership & !cost_burden ~ "noncb_renter",
        !homeownership & cost_burden  ~ "cb_renter",
        TRUE ~ NA_character_),
      
      KITCHEN  = replace(KITCHEN, KITCHEN == 0, NA),
      ROOMS    = replace(ROOMS, ROOMS == 0, NA),
      PLUMBING = replace(PLUMBING, PLUMBING == 0, NA),
  
      severe_housing_problems = if_else(
        KITCHEN == 1 | PLUMBING == 10 | hh_members / ROOMS > 1 | severe_cost_burden, T, F),
      
  #income brackets based on HHINCOME
    #LH-Lower half -> lower40% 
    #UH-Upper Half -> 40th to 80th 
    #T20 -> top 20th percentile
  
      
    income_group = case_when(
      HHINCOME <= quantile(HHINCOME, probs = 0.4) ~ 'LH',
      HHINCOME <= quantile(HHINCOME, probs = 0.8) ~ 'UH',
      HHINCOME > quantile(HHINCOME, probs = 0.8) ~ 'T20',
      TRUE ~ NA_character_),
  
      age_group = case_when(
        age %in% 15:19 ~ NA_character_, 
        age %in% 20:29 ~ "20-29", 
        age %in% 30:39 ~ "30-39",  
        age %in% 40:49 ~ "40-49",  
        age %in% 50:59 ~ "50-59",  
        age %in% 60:69 ~ "60-69", 
        age %in% 70:79 ~ "70-79", 
        age >= 80 ~ "80+")) %>%
    
    group_by(year, SERIAL) %>%
    mutate(
      earner_type = case_when(
        HHINCOME == 0 ~ 'no_earnings',
        any(INCTOT >= 0.95 * HHINCOME) ~ 'single_earner',
        TRUE ~ 'multiple_earner'),
  
      earner_type_d = case_when(
        sex == 'female' & earner_type == 'single_earner' ~ 'single_fem_earner',
        sex == 'male' & earner_type == 'single_earner' ~ 'single_male_earner',
        earner_type == 'multiple_earner' ~ 'multiple_earner'),
        
        num_children = sum(age < 18),
      
        kd_pres = if_else(num_children > 0, "kids", "no_kids"),
      
        own_kd_pres = if_else(NCHILD > 0, "kids", "no_kids")) %>%
    ungroup()
  
  
  census_microdata081122_person %<>% filter(!is.na(age_group))
  
  census_microdata081122 <- census_microdata081122_person %>%
    filter((earner_type == "single_earner" & INCTOT >= 0.95 * HHINCOME) | 
             (earner_type == "multiple_earner" & PERNUM == 1))
  
  arrow::write_feather(census_microdata081122, "census_microdata081122_processed.feather")
  arrow::write_feather(census_microdata081122_person, "census_microdata081122_person_processed.feather")
}

  
```

``` {r louisville 2019 data, include = FALSE}



lville_2019 <- census_microdata081122 %>% 
  filter(FIPS == '21111',
         year == '2019')

cost_burden_age_sf <- census_microdata081122 %>% 
    filter(year %in% 2017:2019) %>%
  mutate(
    cost_burden = factor(cost_burden, 
                         levels = rev(c(TRUE, FALSE)), 
                         labels = rev(c("Cost Burdened", "Non Cost Burdened")), 
                         ordered = TRUE)
    )


```


```{r}


# Inctest = HHINCOME *0.95
# 
# Single_earner_1 = INCTOT >= INCTEST
# single_earner_2 = any(Single_earner_1)

# new_df <- lville_2019 %>%
#   mutate(
#   
#   earner_var = if_else(HHINCOME <= (0.95*INCTOT) , "single_earner", "multi_earner")
#   )
# 
#   
#     earner_var = case_when(
#     HHINCOME == 0 ~ 'no_earnings',
#     INCTOT >= 0.65 * HHINCOME ~ "single_earner",
#     TRUE ~ "multi_earner"
#     )
#   )


      # earner_type = case_when(
      #   HHINCOME == 0 ~ 'no_earnings',
      #   any(INCTOT >= 0.90 * HHINCOME) ~ 'single_earner',
      #   TRUE ~ 'multi_earner'
```


```{r Saves Survey By Demographic Data Frames, include=FALSE, eval=FALSE}

# Data frame with all demographics
# Variable: homeownership
H_gender <- survey_by_demog(census_microdata081122, weight_var = "HHWT", 'homeownership')

# Data frame broken out by earner type rather than male/female
# Variable: homeownership
H_earntype_race <- census_microdata081122 %>%
  group_by(FIPS, year, race, earner_type_d) %>%
  summarize(homeownership = sum(HHWT[homeownership]) / sum(HHWT) * 100, .groups = "drop")

H_earntype  <- census_microdata081122 %>%
  group_by(FIPS, year, earner_type_d) %>%
  summarize(homeownership = sum(HHWT[homeownership]) / sum(HHWT) * 100, .groups = "drop") %>%
  mutate(race = "total")

H_earntype %<>%
  bind_rows(H_earntype_race) %>%
  select(FIPS, year, race, earner_type_d, homeownership)

rm(H_earntype_race)

# Data frame broken out by earner type and presence of children
# Variable: homeownership
H_earntype_kids <- census_microdata081122 %>%
  filter(earner_type == 'single_earner') %>%
  survey_by_demog( weight_var = "HHWT", 'homeownership', other_grouping_vars = c("kd_pres"))

# Data frame broken out by earner type and presence of children
# Variable: cost_burden
CB_earntype_race <- census_microdata081122 %>%
  group_by(FIPS, year, race, earner_type_d) %>%
  summarize(cost_burden = sum(HHWT[cost_burden]) / sum(HHWT) * 100, .groups = "drop")

CB_earntype  <- census_microdata081122 %>%
  group_by(FIPS, year, earner_type_d) %>%
  summarize(cost_burden = sum(HHWT[cost_burden]) / sum(HHWT) * 100, .groups = "drop") %>%
  mutate(race = "total")

CB_earntype %<>%
  bind_rows(CB_earntype_race) %>%
  select(FIPS, year, race, earner_type_d, cost_burden)

rm(CB_earntype_race)

save(H_gender, H_earntype, H_earntype_kids, CB_earntype, 
     file = "clean_svybydemog_data.RData")
```

```{r}
load("clean_svybydemog_data.RData")
```

Gender equity in Louisville is an unfortunately unequal reality. Women are worse off in key standard of living areas such as household income and homeownership. Additionally, these issues are exasperated for women from a one-income home, women with children, and minority women. Disproportionate cost of living burdens and care-taking responsibilities can perpetuate a viscous cycle of inequity. Understanding the true size of the 'equity gap' can help inform policy decisions to stop this cycle from continuing.  

Key definitions:

* **Cost-burdened household** - when a household has to pay more than *30%* of their income toward housing costs (including rent, utilities, mortgage payments, and any other homeownership costs.)
* **Severely cost-burdened household** - when a household has to pay more than *50%* of their income toward housing costs.
* **Single-income household** and **Multiple-earner household** - A single-income household has only one wage earner, while a multiple-earner household includes multiple earners. Throughout the data, we usually group households into three groups:
  + Multiple-earner households
  + Single-earner, female-headed households
  + Single-earner, male-headed households


Key Takeaways:

1. Women and men in single-income households have very similar rates of homeownership in Louisville. 
2. On average, women in single-income households make significantly less money on average than men from single-income households. As a result, almost half of Louisville women in single-income households are cost-burdened, putting them at increased risk of eviction or foreclosure.
3. Around 43% of women in single-income households earn a living wage that covers their basic expenses (around \$30,000.) However, that number is only around 10% for single-income women with a child, only around 5% for single-income women with two children, and near 0% for single-income women with three or more children.
4. Compared to peer cities, Louisville has relatively high homeownership for women in single-income households with no children. However, for single-income women with children, we are second to last. Additionally, homeownership for women from a one-income household with children has been steadily decreasing since 2016. 
5. Homeownership for single-income women is much lower for women of color.


Rolling mean explanation

Race categories explanation

GLP strives to use inclusive language and analyze data for traditionally underrepresented groups whenever possible. However, current data has its limitations. The terminology we use to describe race, sex, gender, and other identities mirrors the way questions were asked in the U.S. Census Bureau's American Community Survey. Additionally, the survey does not provide us with enough information to create data on many populations in Louisville. When we break data down by race, we include data for white non-Hispanic residents, Black non-Hispanic residents, and Hispanic residents.




Data are often scarce for Hispanic and Latinx populations, as well for the LGBTQ+
population. National data we collect–for measures such as overdoses and certificates–are often unavailable at the zip code or neighborhood levels. 


# Waffle Chart
```{r Harrison, out.width = "100%", out.extra='style="border: 0px" allowfullscreen=""'}

waffle_data <- survey_by_demog(lville_2019, "hh_type", other_grouping_vars = c("kd_pres", "earner_type"))

gender <- waffle_data %>%
  filter(
    race == "total",
    sex != "total",
    var_type == "population") %>%
  group_by(sex) %>%
  summarize(n = sum(cb_homeowner), .groups = "drop") %>%
  mutate(pct = n / sum(n) * 100)

earn_type <- waffle_data %>%
  filter(
    race == "total",
    sex != "total",
    var_type == "population") %>%
  group_by(sex, earner_type)  %>%
  summarize(n = sum(cb_homeowner), .groups = "drop") %>%
  mutate(pct = n / sum(n) * 100)

earn_type <- waffle_data %>%
  filter(
    race == "total",
    sex != "total",
    var_type == "estimate") %>%
  group_by(sex, earner_type)  %>%
  summarize(n = sum(cb_homeowner + noncb_homeowner),
            total = sum(cb_homeowner + noncb_homeowner + cb_renter + noncb_renter),
            .groups = "drop") %>%
  mutate(pct = n / sum(total) * 100)

earn_type <- waffle_data %>%
  filter(
    race == "total",
    sex != "total",
    var_type == "estimate") %>%
  group_by(sex, earner_type)  %>%
  summarize(n = sum(cb_homeowner + cb_renter),
            total = sum(cb_homeowner + noncb_homeowner + cb_renter + noncb_renter),
            .groups = "drop") %>%
  mutate(pct = n / sum(total) * 100)
  
demog_data <- data.frame(
  row_num = rep(1:10, each = 10),
  col_num = rep(10:1, 10),
  gender = c(rep("male", 52), rep("female", 48)),
  earn_type = c(rep("multi_earn", 28), rep("single_earn", 20), rep("multi_earn", 29), rep("single_earn", 23)))



knitr::include_url("https://greaterlou.shinyapps.io/gender_demogs/", height = "2500px")


```

```{r, eval=FALSE}

# Single Earner Vs. Multiple Earner Women

t_df <- census_microdata081122 %>%
  filter(sex == "female")

prop.table(table(t_df$earner_type_d))

#single earner vs multi earner who women who happen to be head of household
#should we remove head of household filter?
```


# Income {.tabset}

```{r, include = FALSE}
#Finding income ranges

#95th perc of income
cut_95 = quantile(lville_2019$HHINCOME, probs = 0.95)

```

This section will display income levels for single earner males, females, and multiple demographic and household makeups for single females earners. 

Disparity between single earner male and female wages is significant in Louisville. In 2019, 50% of males in Louisville made at least \$45,000 while this values is only \$30,000 for females. This disparity increases as wages rise, wealthiest 10% of males earn at least \$120,000 while females in the same income bracket earn at least \$71,000. 
```{r 2019 Louisville Men and Women Income Compared, eval=FALSE}

#fix formatting
single_earner_pctiles <- lville_2019 %>%
  group_by(sex) %>%
  summarize(
    ten_pct = Hmisc::wtd.quantile(HHINCOME, HHWT, probs = 0.1),
    twenty_five_pct = Hmisc::wtd.quantile(HHINCOME, HHWT, probs = 0.25),
    fifty_pct = Hmisc::wtd.quantile(HHINCOME, HHWT, probs = 0.5),
    seventy_five_pct = Hmisc::wtd.quantile(HHINCOME, HHWT, probs = 0.75),
    ninety_pct = Hmisc::wtd.quantile(HHINCOME, HHWT, probs = 0.9))

library(gt) 

gt(single_earner_pctiles) %>%
  tab_header(title = "Income Percentiles by Sex",
             subtitle = "") %>%
  fmt_currency(columns = vars(ten_pct, twenty_five_pct, fifty_pct, seventy_five_pct,
                              ninety_pct),
                            use_subunits = F) %>%
  cols_label(ten_pct = "10th",
             twenty_five_pct = "25th",
             fifty_pct = "Median",
             seventy_five_pct	= "75th",
             ninety_pct = "90th") %>%
  cols_align(align = "center") %>%
      tab_source_note(
    source_note = md("Source: ACS microdata from IPUMS-USA")) %>%
  opt_row_striping(row_striping = TRUE) %>%
  opt_table_outline() %>%
  tab_options(
    table.font.size = px(12),
    table.width = pct(50)) %>%
  tab_style(
    cell_text(
      font = "Montserrat",
      weight = "bold"), 
    cells_row_groups())
```


## by Gender

```{r}
p <- lville_2019 %>% 
  filter(HHINCOME <= cut_95,
         earner_type == "single_earner") %>%
  func_plt_hist_overlay( "sex")

p <- p + glp_graph_theme

p <- p + labs(
  title = "Single Earner Income by Gender",
) +
  
  ylab(" ") +
  
  guides(color = FALSE) + 
  
  facet_wrap(~sex, nrow = 2) +
  
  theme(
  #axis.ticks.x =  element_line(size = 50000),
  strip.text = element_blank()

  )  +

scale_x_continuous(
  breaks = c(50000, 100000, 150000, 200000),
  label = c("$50k", "$100k", "$150k", "$200k")
) +
  scale_y_continuous(labels = scales::comma) +
  
  scale_color_manual(values=c("#f58021", "#00a9b7"))+
  
  scale_fill_manual(
    values = c("#f58021", "#00a9b7"), 
    labels = c("Female", "Male"))

p
```

```{r, eval= FALSE}

### Percent
temp_df <- lville_2019 %>% 
  filter(HHINCOME <= cut_95,
         earner_type == "single_earner")

  p_percent <- ggplot(temp_df, aes(x=HHINCOME, 
                                   y = (..count..)/sum(..count..),
                                   fill=sex, 
                                   color = sex, 
                                   weight = HHWT)) +
    geom_histogram(alpha=0.5, position = 'identity', binwidth = 10000) +
    scale_fill_manual(values = c("#00A9B7", "#F58021")) +
    scale_color_manual(values = c("#00A9B7", "#F58021")) +

    labs(fill="") +
    xlab("Household Income") +
    ylab("Percentage") 

p_percent <- p_percent + glp_graph_theme

p_percent <- p_percent + labs(
  title = "Single Earner Income by Gender",
) +
  
  ylab(" ") +
  
  guides(color = FALSE) + 
  
  facet_wrap(~sex, nrow = 2) +
  
  theme(
  #axis.ticks.x =  element_line(size = 50000),
  strip.text = element_blank()

  )  +

scale_x_continuous(
  breaks = c(50000, 100000, 150000, 200000),
  label = c("$50k", "$100k", "$150k", "$200k")
) +
  scale_y_continuous(labels=percent)

p_percent
```

## by Age

-this seems wrong...shouldn't the groups be flippled logically?...also there arent any multiple earners if you examine the dataframe
Single and multiple income households follow a similar pattern of earnings with highest earnings seen between 40 and 49 years of age. 

```{r Median Income for Earning Group by Age}

I_median_earn_age <- lville_2019 %>%
  group_by(age_group, earner_type_d) %>%
  summarize(Med=median(HHINCOME)) 

I_median_earn_age_plot <- ggplot(I_median_earn_age, 
       aes(x=age_group, y=Med, fill = earner_type_d)) + 
  geom_bar(stat="identity", position='dodge') 

I_median_earn_age_plot <- I_median_earn_age_plot + glp_graph_theme

I_median_earn_age_plot <- I_median_earn_age_plot +
  labs(
  title = "Median Earnings by Age Group",
) + 
  ylab("Household Income") +
  xlab("Age Group") +
  
  scale_y_continuous(labels = scales::dollar) +   
  scale_fill_manual(
    values = c("#0E4A99", "#F58021"), 
    labels = c("Multiple Earner", "Single Female Earner"))

  
I_median_earn_age_plot

```

## by Race {.tabset}

When examining single earner females broken down by race it becomes apparent that for non-white females it is significantly less likely to see wages about \$50,000. 

Only wages under \$150,000 are displayed because of the small number of instances any female single earners make over $150,000.

```{r, include=FALSE}

func_income_by_race <- function(race_group) {



  q <- census_microdata081122 %>% 
  filter(
    FIPS == "21111",
    year %in% 2016:2019,
    sex == 'female',
    race == race_group,
    earner_type == 'single_earner',
    HHINCOME <= cut_95) 
  
   q <- q %>%
   ggplot( aes(x=HHINCOME, 
              y = (..count..)/sum(..count..),
               fill=race, 
               color = race, 
               weight = HHWT)) +
    geom_histogram(alpha=0.5, position = 'identity', binwidth = 10000) 


q <- q + glp_graph_theme

q <- q + 
  labs(
  title = "Black Female Single Earner Income",
) + 
  ylab(" ") +
  xlab("Household Income")+
  
  guides(color = FALSE)

q <- q + 
  
  theme( 
  #axis.ticks.x =  element_line(size = 50000),
  strip.text = element_blank()
  
  )  + 
  
scale_x_continuous(
  breaks = c(50000, 100000, 150000),
  label = c("$50k", "$100k", "$150k")
) +
  scale_y_continuous(labels = scales::percent)

    
  
  return (q)
  
}
```

```{r, eval=FALSE}

## original faceted graph...has been removed
sing_fem_inc_race<-census_microdata081122 %>% 
  filter(
    FIPS == "21111",
    year %in% 2016:2019,
    sex == 'female',
    earner_type == 'single_earner',
    HHINCOME <= cut_95) 
  
   sing_fem_inc_race_plt <- sing_fem_inc_race %>%
   ggplot( aes(x=HHINCOME, 
              y = (..count..)/sum(..count..),
               fill=race, 
               color = race, 
               weight = HHWT)) +
    geom_histogram(alpha=0.5, position = 'identity', binwidth = 10000) 


sing_fem_inc_race_plt <- sing_fem_inc_race_plt + facet_wrap(~race, nrow = 2) 

sing_fem_inc_race_plt <- sing_fem_inc_race_plt + glp_graph_theme

sing_fem_inc_race_plt <- sing_fem_inc_race_plt + 
  labs(
  title = "Female Single Earner Income",
) + 
  ylab(" ") +
  xlab("Household Income")
  
  # guides(color = FALSE)

sing_fem_inc_race_plt <- sing_fem_inc_race_plt + 
  
  theme( 
  #axis.ticks.x =  element_line(size = 50000),
  strip.text = element_blank()
  
  )  + 

scale_x_continuous(
  breaks = c(50000, 100000, 150000),
  label = c("$50k", "$100k", "$150k")
) +
  scale_y_continuous(labels = scales::percent)

sing_fem_inc_race_plt <- sing_fem_inc_race_plt +
  
  scale_fill_manual(values = c("#0E4A99", "#F58021","#00A9B7", "#800055")) +
  scale_color_manual(values = c("#0E4A99","#F58021","#00A9B7", "#800055")) 

sing_fem_inc_race_plt
```


### Black
```{r 2019 Louisville Single Female Income by Race}

black_female_earner <- func_income_by_race("black")
black_female_earner

```

### Hispanic
```{r}

hisp_female_earner <- func_income_by_race("hispanic")
hisp_female_earner <- hisp_female_earner + 
  labs(
  title = "Hispanic Female Single Earner Income",
) + 
  scale_fill_manual(values = "#0E4A99") +
  scale_color_manual(values = "#0E4A99")
  
hisp_female_earner
```


### White
```{r}
white_female_earner <- func_income_by_race("white")
white_female_earner <- white_female_earner + 
  labs(
  title = "White Female Single Earner Income",
)  + 
  scale_fill_manual(values = "#F58021") +
  scale_color_manual(values = "#F58021")

white_female_earner
```

### Other
```{r}
other_female_earner <- func_income_by_race("other")
other_female_earner <- other_female_earner + 
  labs(
  title = "Other Female Single Earner Income",
) + 
  scale_fill_manual(values = "#00A9B7") +
  scale_color_manual(values = "#00A9B7")

other_female_earner
```

## Compared to a Living Wage {.tabset}

For single-adult households, a living wage is \$33k for a single adult, \$66k for an adult with one child, \$84k for an adult with two children, and $112k for one adult and three children.

```{r}
func_income_by_kids <- function(num_kids, living_wage) {



  w <- census_microdata081122 %>% 
  filter(
    FIPS == "21111",
    year %in% 2016:2019,
    sex == 'female',
    NCHILD == num_kids,
    earner_type == 'single_earner',
    HHINCOME <= cut_95) 
  
   w <- w %>%
   ggplot( aes(x=HHINCOME, 
              y = (..count..)/sum(..count..),
              fill = sex,
              group = sex,
               weight = HHWT)) +
    geom_histogram(alpha=0.5, position = 'identity', binwidth = 10000) +
     geom_vline( aes(xintercept = living_wage), linetype = "dashed",  colour="blue", size = 1.5) 
   

#sing_fem_inc_race_plt <- sing_fem_inc_race_plt + facet_wrap(~race, nrow = 2) 

w <- w + glp_graph_theme

w <- w + 
  labs(
  title = "Black Female Single Earner Income",
) + 
  ylab(" ") +
  xlab("Household Income")+
  
  guides(color = FALSE)

w <- w + 
  
  theme( 
  #axis.ticks.x =  element_line(size = 50000),
  strip.text = element_blank()
  
  )  + 
  
scale_x_continuous(
  breaks = c(50000, 100000, 150000),
  label = c("$50k", "$100k", "$150k")
) +
  scale_y_continuous(labels = scales::percent) 
    
  
  return (w)

}
```

### No Child

```{r}


#why is color not working?
#still need to add living wage lines

under_liv_wage_0 <- census_microdata081122 %>% 
  filter(
    FIPS == "21111",
    year %in% 2016:2019,
    sex == 'female',
    NCHILD == 0,
    earner_type == 'single_earner') %>%
  group_by(HHINCOME < 33321.6) %>%
  summarize(count = sum(HHWT)) #a little more than half are earning a living wage
#do this for each graphof this type...add info above chunk

no_kids_female_earner <- func_income_by_kids(0, 33321.6)
no_kids_female_earner <- no_kids_female_earner + 
  labs(
  title = "Female Single Earner Income, No Children",
) + 
  scale_fill_discrete(labels = "No Children")

no_kids_female_earner
```

### 1 Child
```{r}
under_liv_wage_1 <- census_microdata081122 %>% 
  filter(
    FIPS == "21111",
    year %in% 2016:2019,
    sex == 'female',
    NCHILD == 1,
    earner_type == 'single_earner') %>%
  group_by(HHINCOME < 66081.6) %>%
  summarize(count = sum(HHWT))

one_child <- func_income_by_kids(1, 66081.6)
one_child <- one_child + 
  labs(
  title = "Female Single Earner Income, One Child",
) + 
  scale_fill_manual(values = "#800055", labels = "One Child" ) +
  scale_color_manual(values = "#800055") 

one_child
```


### 2 Children
```{r}

under_liv_wage_2 <- census_microdata081122 %>% 
  filter(
    FIPS == "21111",
    year %in% 2016:2019,
    sex == 'female',
    NCHILD == 2,
    earner_type == 'single_earner') %>%
  group_by(HHINCOME < 83990.4) %>%
  summarize(count = sum(HHWT))

two_child <- func_income_by_kids(2, 83990.4)
two_child <- two_child + 
  labs(
  title = "Female Single Earner Income, Two Children",
) + 
  scale_fill_manual(values = "#356E39", labels = "Two Children") +
  scale_color_manual(values = "#356E39") 

two_child
```

### 3 Children

```{r}
under_liv_wage_3 <- census_microdata081122 %>% 
  filter(
    FIPS == "21111",
    year %in% 2016:2019,
    sex == 'female',
    NCHILD == 3,
    earner_type == 'single_earner') %>%
  group_by(HHINCOME < 111529.6) %>%
  summarize(count = sum(HHWT))

three_child <- func_income_by_kids(3, 111529.6)
three_child <- three_child + 
  labs(
  title = "Female Single Earner Income With Three Children",
) + 
  scale_fill_manual(values = "#CFB94C", labels = "Three Children") +
  scale_color_manual(values = "#CFB94C") 

three_child
```


```{r}
## Avg Age of Women:Living Wage

age_liv_wage <- function(num_kids, living_wage) {

df <- census_microdata081122 %>% 
  filter(
    FIPS == "21111",
    year %in% 2016:2019,
    sex == 'female',
    NCHILD == num_kids,
    earner_type == 'single_earner',
    HHINCOME < living_wage) %>%
  #group_by(HHINCOME < living_wage) %>%
  summarize(Hmisc::wtd.mean(age, HHWT, na.rm=TRUE))

return(df)

}

# three_child <- func_income_by_kids(3, 101452.61)


under_liv_wage_0_age <- age_liv_wage(0, 30303.98) #61.2
under_liv_wage_1_age <- age_liv_wage(1, 60264.75) #38.9
under_liv_wage_2_age <- age_liv_wage(2, 76451.81) #34.9
under_liv_wage_3_age <- age_liv_wage(3, 101452.61) #32.7

```

# Cost Burden {.tabset}

Add cost burden definition reminder?


### by Income

Visualizing cost burden status by household income provides a clear indication that very few single female earners are cost burdened when earning at least \$70,000. For reference, %50 of single female earners in Louisville make less than \$30,000.
```{r}


these_labels <- paste0(dollar(seq(1, 273500, 10000), scale = 0.001, accuracy = 1, suffix = "k"))

cost_burden_sf <- lville_2019 %>% 
  filter(
    sex == 'female',
    earner_type == 'single_earner',
    HHINCOME <= cut_95) %>%
  mutate(
    cost_burden = factor(cost_burden, 
                         levels = rev(c(TRUE, FALSE)), 
                         labels = rev(c("Cost Burdened", "Non Cost Burdened")), 
                         ordered = TRUE),
    inc_bins = cut(HHINCOME, seq(1, 283500, 10000),
                   labels = these_labels) %>%
      factor(levels = these_labels, ordered = TRUE)
    )
    
temp_df <- cost_burden_sf %>%
  group_by(inc_bins, cost_burden) %>%
  summarize(count = sum(HHWT), .groups = "drop") %>%
  complete(inc_bins, cost_burden, fill = list(count = 0)) %>%
  filter(!is.na(inc_bins)) %>%
  group_by(inc_bins) %>%
  mutate(percent = count / sum(count)) %>%
  ungroup() %>%
  filter(cost_burden == "Cost Burdened")
 
temp_df <- temp_df[1:14,]

cost_burden_sf_plot <- ggplot(temp_df, 
       aes(x = inc_bins,
           y = percent,
           group = 1)) +
  geom_line(linetype = "dashed",  color="chartreuse4", size=3) +
  geom_point(color="chartreuse4", size=8)

cost_burden_sf_plot <- cost_burden_sf_plot + glp_graph_theme

cost_burden_sf_plot <- cost_burden_sf_plot + 
  labs(
  title = "Female Single Earner Cost Burden Level by Income",
) + 
  ylab(" ") +
  xlab("Household Income") +
  
  guides(color = FALSE) +
  
  theme(
  strip.text = element_blank()

  )  +

  scale_y_continuous(labels = scales::percent)

cost_burden_sf_plot
```

### by Age 

While cost burden status does appear to decrease on average for single female earners as their age increases, there is significant room for improvement. Notice that a significant portion of individuals in their twenties face cost burden status and after single female earners reach their 60's cost burden status begins to increase again.  
```{r  Cost Burdened Stage by Age/Race}
cost_burden_age_sf %<>% drop_na(cost_burden) #this will need to be run once and then left alone if tweaking graphs

temp_df1 <- cost_burden_age_sf %>%
  filter(earner_type_d == "single_fem_earner") %>%
  mutate(
    age_group = case_when(
      age %in% 15:19 ~ NA_character_, 
      age %in% 20:29 ~ "20-29", 
      age %in% 30:39 ~ "30-39",  
      age %in% 40:49 ~ "40-49",  
      age %in% 50:59 ~ "50-59",  
      age %in% 60:69 ~ "60-69", 
      age %in% 70:79 ~ "70-79", 
      age >= 80 ~ "80+"))

temp_df1 %<>%  filter(!is.na(age_group))

cost_burden_age_sf_facet_plt <- ggplot(temp_df1,
       aes(x=age_group, y=HHWT , fill=cost_burden),
        color="#00A9B7") +
geom_bar(stat="identity", position='fill')
  
#facet_wrap(~earner_type_d)

cost_burden_age_sf_facet_plt <- cost_burden_age_sf_facet_plt + glp_graph_theme

cost_burden_age_sf_facet_plt <- cost_burden_age_sf_facet_plt + 
    theme(
    legend.position = "right",
    
    strip.text = element_text(size = 40)
    ) +
  labs(
  title = "Cost Burdened Status by Age"
) + 
  ylab(" ") +
  xlab(" ") +
  scale_fill_discrete(labels = c("Non Cost Burdened", "Cost Burdened")) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +

  scale_y_continuous(labels = scales::percent)

cost_burden_age_sf_facet_plt





```

## Over Time

While Louisville does fair better than its peer cities with less cost burdened single female earners, there is significant room for improvement in terms of closing the gap between single female and male earners. 
```{r Cost-burdened by earner type Trend }

CB_earntype %<>%
  filter(
    race == 'total') %>%
  select( -race) %>%
  pivot_wider(names_from = "earner_type_d", values_from = "cost_burden")
  

trend(CB_earntype, 
      multiple_earner:single_fem_earner:single_male_earner, 
      pctiles = F,
      plot_title = "Cost Burden by Earner Type",
      cat = c("Multiple Earners" = "multiple_earner", "Single Female Earner" = "single_fem_earner", "Single Male Earner" = "single_male_earner"),
      y_title = 'Percent',
      caption_text = 
      "Source Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")



```

```{r, eval = FALSE}
m <- trend_data_maxmin(CB_earntype, "single_fem_earner")

# creates change variable starting at year 2000...perc point change in female sing eaners that are cost burdened
CB_earntype %>%
group_by(FIPS) %>%
mutate(change = single_fem_earner - single_fem_earner[year == 2000]) %>%
ungroup()%>%
  ranking_data("change")%>% 
  pull_peers(add_info=T)
```


# Homeownership {.tabset}
Overall, 32% of homeowners in Louisville are female with and average age of 64. - ask harrison about this

## by demographic

```{r, eval=FALSE}
### Female Homeowner Proportions

w_howner <- lville_2019 %>%
  filter(sex == "female")

prop.table(table(w_howner$homeownership)) 
```

```{r, eval = FALSE}
#all homeownership

prop.table(table(lville_2019$homeownership, lville_2019$sex)) 
```

```{r, eval = FALSE}
### Age of Female Homeowners

female_age_howner <- lville_2019 %>%
  filter(sex == "female") %>%
  group_by(homeownership) %>%
  summarize(
    ten_pct = Hmisc::wtd.quantile(age, HHWT, probs = 0.1),
    twenty_five_pct = Hmisc::wtd.quantile(age, HHWT, probs = 0.25),
    fifty_pct = Hmisc::wtd.quantile(age, HHWT, probs = 0.5),
    seventy_five_pct = Hmisc::wtd.quantile(age, HHWT, probs = 0.75),
    ninety_pct = Hmisc::wtd.quantile(age, HHWT, probs = 0.9),
    avg = mean(age))

female_age_howner #avg age for female homeowner is 64.06362

```


```{r, eval = FALSE}

### Female Homeowners with & withut Kids Proportions

kid_howner <- lville_2019 %>%
  filter(sex == "female",
         homeownership == TRUE)

prop.table(table(kid_howner$kd_pres)) #0.3379953 women homeowners have kids

```


It is clear that having multiple household incomes significantly increases the ability to own a home. On average, single female home ownership has declined in Louisville since 2007. 
```{r Trend in Homeownership by Gender}

#also compares earner types
temp_df <- H_earntype %>%
  filter(race == 'total') %>%
  pivot_wider(names_from = "earner_type_d", values_from = "homeownership")

trend(temp_df, 
      multiple_earner:single_male_earner,
      plot_title = "Homeownership by Year", 
      cat = c("Multiple Earners" = "multiple_earner", "Single Female" = "single_fem_earner", "Single Male" = "single_male_earner"),
      pctiles = F,
      y_title = 'Percent',
      rollmean = 3,
      caption_text = 
      "Source: Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```

## by presence of children {.tabset}

### trend

For women in Louisville with children home ownership rates are drastically lower than rates for women without children. Additionally, the female home onwership rate with children has steadily decreased since 2014. 
```{r Trend in Homeownership by for Single Females with vs. without Kids}

H_earntype_kids %<>%
  filter(
    var_type == 'percent',
    sex == "female") %>%
  pivot_wider(names_from = 'kd_pres', values_from = 'homeownership') %>%
  select(-sex)


trend(H_earntype_kids, 
      kids:no_kids,
      rollmean = 3,
      plot_title = "Female Homeownership by Presence of Children", 
      cat = c("Children" = "kids", "No Children" = "no_kids"), 
      y_title = 'Percent',
      caption_text = 
      "Source: Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```

### Without children

```{r Ranking Single Female Homeownership}
ranking(H_earntype_kids, 
        'no_kids',
        plot_title = "Single Earner Female Homeownership",
        caption_text = 
      "Source: Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```


### Ranking with Children

```{r Ranking Single Female with Kids Homeownership}
ranking(H_earntype_kids,
        'kids',
        plot_title = "Single Earner Female Homeownership with Children",
        #title_scale = 0.8,
        caption_text = 
      "Source: Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```

## by Race

There is a significant disparity in home ownership rates for non-white female single earners. Black single female earners have seen a steady decrease since 2007. 
```{r Trend in Single Female Homeownership by Race}
trend(filter(H_earntype, earner_type_d == "single_fem_earner"), 
      homeownership, 
      rollmean = 5,
      pctiles = F,
      plot_title = "Single Female Homeownership by Year", 
      cat = 'race', 
      y_title = 'Percent',
      caption_text = 
      "Source Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```

## by Race and Children {.tabset}

A similar trend to the overall average is scene for each demographic when comparing home ownership rates for females with and without children - rates are lower for females with children. 
### With Children

```{r Trend in Single Female Homeownership by Race with and without kids}

trend(H_earntype_kids, 
      kids, 
      rollmean = 5,
      pctiles = F,
      plot_title = "Single Female Homeownership by Year with Children", 
      cat = 'race', 
      y_title = 'Percent',
      caption_text = 
      "Source Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```

### Without Children

```{r}
trend(H_earntype_kids, 
      no_kids, 
      rollmean = 5,
      pctiles = F,
      plot_title = "Single Female Homeownership by Year without Children", 
      cat = 'race', 
      y_title = 'Percent',
      caption_text = 
      "Source Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")
```


```{r, include = FALSE, eval=FALSE}

## Affordable Housing need for Female Single Earners

#variable of interest
#cost_burden
#severe_cost_burden
#severe_housing_problems

#fiter to single women - done
#single women total - 72032
#single women are cost burdened - 34683
#single women are severley cost burdened - 17028
#single women have severe housing problems - 17935
#group by cost burden and severe housing problems and find values - 17357
  #notable to conclude A.our def of housing problems and most ppl face these because of low income
  #

val <- lville_2019 %>%
  filter(
    sex == "female",
    earner_type == "single_earner"
  ) %>%
  group_by(severe_housing_problems, cost_burden) %>%
  summarize(n=sum(HHWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n)
  ) 
#Cost Burden: 81655
#Severe Cost Burden: 38830
#Severe Housing Problems: 47392

#use these stats in a paragraph

val
```



# Education

## Education Level {.tabset}

### By Gender

Compared to single male earners, single female earners see a smaller percentage of individuals with High School Degree but larger percentages of individuals with Bachelor or Graduate degrees.  

```{r}
E_singM_singF <- census_microdata081122_person %>% 
  filter(year %in% 2017:2019, 
         earner_type == 'single_earner') %>%
  group_by(sex, educ) %>%
  summarize(n=sum(PERWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100,
    educ = factor(educ, 
                  levels = rev(c("no_hs", "hs", "some_col", "assoc", "bach","grad")), 
                  ordered = TRUE))


E_singM_singF_plot <- ggplot(E_singM_singF, 
       aes(x=sex, 
           y=rate, 
           fill = educ)) + 
  geom_bar(stat="identity", position = "fill") 


#E_singM_singF_plot <- E_singM_singF_plot + facet_wrap(~kd_pres) 
  
E_singM_singF_plot <- E_singM_singF_plot + glp_graph_theme

E_singM_singF_plot <- E_singM_singF_plot + 
  
  theme(
    legend.position = "right"
    ) +

  labs(
  title = "Single Earner Education Levels by Gender",
) + 
  scale_fill_discrete(
    labels = c("Graduate","Bachelor", "Associate", "Some College",  "High School", "No High School")) + 
  scale_x_discrete (labels = c("female" = "Female", "male" = "Male")) +
  
  scale_y_continuous(labels = scales::percent)

E_singM_singF_plot
```


### By presence of Children

--because they are so similar can we make this just for females? 
```{r}


E_singM_singF <- census_microdata081122_person %>% 
  filter(year %in% 2017:2019, 
         earner_type == 'single_earner') %>%
  group_by(sex, educ, kd_pres) %>%
  summarize(n=sum(PERWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100,
    educ = factor(educ, 
                  levels = rev(c("no_hs", "hs", "some_col", "assoc", "bach","grad")), 
                  ordered = TRUE))


E_singM_singF_plot <- ggplot(E_singM_singF, 
       aes(x=sex, 
           y=rate, 
           fill = educ)) + 
  geom_bar(stat="identity", position = "fill") 


E_singM_singF_plot <- E_singM_singF_plot + facet_wrap(~kd_pres) 
  
E_singM_singF_plot <- E_singM_singF_plot + glp_graph_theme

E_singM_singF_plot <- E_singM_singF_plot + 
  
  theme(
    legend.position = "right"
    ) +

  labs(
  title = "Single Earner Education Levels by Gender and Children",
) + 
  ylab(" ") +
  xlab(" ") +
  scale_fill_discrete(
    labels = c("Graduate","Bachelor", "Associate", "Some College",  "High School", "No High School")) + 
  scale_x_discrete (labels = c("female" = "Female", "male" = "Male")) +
  
  scale_y_continuous(labels = scales::percent)

E_singM_singF_plot
```

### by race
```{r, pie chart}

E_singF_race <- lville_2019 %>% 
  filter(
    sex == 'female',
    earner_type == 'single_earner') %>%
  group_by(race, educ) %>%
  summarize(n=sum(PERWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100,
    educ = factor(educ, 
                  levels = rev(c("no_hs", "hs", "some_col", "assoc", "bach","grad")), 
                  ordered = TRUE))

ggplot(E_singF_race, aes(x = "", y = rate, fill = educ)) +
  geom_bar(stat="identity", width=2, color="white") +
  geom_text(aes(label = scales::percent(rate, scale = 1, accuracy = 0.1)),
            position = position_stack(vjust = 0.5),
            size = 10,
            color = "white") +
  coord_polar("y", start=0) +
  
    glp_graph_theme + 
  
  scale_fill_manual(name = "",
                    values = c("#0E4A99", "#F58021", "#00A9B7", "#800055", "#356E39", "#CFB94C"),
                    labels = c("Graduate","Bachelor", "Associate", "Some College",  "High School", "No High School")) +
  
  labs(
    title = "") + 
  
    theme(
  axis.title.x=element_blank(),
  axis.title.y=element_blank(),
  axis.text = element_blank()
  #strip.text = element_blank()

  )  +

  facet_wrap(~race) 

```



```{r, edu by race and kids, eval=FALSE}

### none of the edu below will be used
edu_by_kids <- function(race_group) {
  
  v <- census_microdata081122_person %>% 
  filter(year %in% 2017:2019, 
         earner_type == 'single_earner',
         race == race_group) %>%
  group_by(sex, educ, kd_pres) %>%
  summarize(n=sum(PERWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100,
    educ = factor(educ, 
                  levels = rev(c("no_hs", "hs", "some_col", "assoc", "bach","grad")), 
                  ordered = TRUE))


v <- ggplot(E_singM_singF, 
       aes(x=sex, 
           y=rate, 
           fill = educ)) + 
  geom_bar(stat="identity", position = "fill") 


v <- v + facet_wrap(~kd_pres) 
  
v <- v + glp_graph_theme

v <- v + 
  
  theme(
    legend.position = "right"
    ) +

  labs(
  title = "Single Earner Education Levels by Gender",
) + 
  ylab(" ") +
  xlab(" ") +
  scale_fill_discrete(
    labels = c("Graduate","Bachelor", "Associate", "Some College",  "High School", "No High School")) + 
  scale_x_discrete (labels = c("female" = "Female", "male" = "Male")) +
  
  scale_y_continuous(labels = scales::percent)

v
  
}
```


```{r, eval = FALSE}
edu_by_kids("black")

edu_by_kids("white")


edu_by_kids("hispanic")

edu_by_kids("other")

```


```{r, Education of Single Female by race, eval = FALSE}


E_singF_race <- lville_2019 %>% 
  filter(
    sex == 'female',
    earner_type == 'single_earner') %>%
  group_by(race, educ) %>%
  summarize(n=sum(PERWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100,
    educ = factor(educ, 
                  levels = rev(c("no_hs", "hs", "some_col", "assoc", "bach","grad")), 
                  ordered = TRUE))

E_singF_race_plot <- ggplot(E_singF_race, aes(x=race, y=rate, fill=educ)) + 
geom_bar(stat="identity", position='fill')


E_singF_race_plot <- E_singF_race_plot + glp_graph_theme

E_singF_race_plot <- E_singF_race_plot + 
    theme(
    legend.position = "right"
    ) +
  labs(
  title = "Single Female Education Breakdown",
) + 
  ylab(" ") +
  xlab("Race") +
  scale_fill_discrete(labels = c("Graduate","Bachelor", "Associate", "Some College",  "High School", "No High School")) + 
  scale_x_discrete (labels = c("female" = "Female", "male" = "Male")) +
  
  scale_y_continuous(labels = scales::percent)

E_singF_race_plot


```



```{r Proportion of Earner Types over time, eval=FALSE}

##earner types over time
earner_trend <- census_microdata081122_person %>%
  
  group_by(year, earner_type_d) %>%
  summarize(n=sum(HHWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100) 
  
earner_trend_plt <- ggplot(earner_trend, 
       aes(x=year, y=rate, fill=earner_type_d),
        color="#00A9B7") + 
geom_bar(stat="identity", position='fill')


earner_trend_plt <- earner_trend_plt + glp_graph_theme

earner_trend_plt <- earner_trend_plt + 
    theme(
    legend.position = "right"
    #strip.text = element_blank()
    ) +
  labs(
  title = "Earner Type Trend"
) + 
  ylab(" ") +
  xlab(" ") +
  scale_fill_discrete(labels = c("Multiple Earner", "Single Female Earner", "Single Male Earner")) + 

  scale_y_continuous(labels = scales::percent)

earner_trend_plt
```

# Appendix {.tabset}

## Gender

>What is Person X's sex? Mark (X) ONE box.
>
>[ ] Male
>
>[ ] Female

## Hispanic Descent

>Is Person X of Hispanic, Latino, or Spanish origin?
>
>[ ] No, not of Hispanic, Latino, or Spanish origin
>
>[ ] Yes, Mexican, Mexican Am., Chicano
>
>[ ] Yes, Puerto Rican
>
>[ ] Yes, Cuban
>
>[ ] Yes, another Hispanic, Latino, or Spanish origin -- Print origin, for example, Argentinean, Colombian, Dominican, Nicaraguan, Salvadoran, Spaniard, and so on. --> ______________________________________

## Race

>What is Person X's race? Mark (X) one or more boxes.
>
>[ ] White
>
>[ ] Black or African Am.
>
>[ ] American Indian or Alaska Native -- Print name of enrolled or principal tribe. --> __________________
>
>[ ] Asian Indian
>
>[ ] Japanese
>
>[ ] Native Hawaiian
>
>[ ] Chinese
>
>[ ] Korean
>
>[ ] Guamanian or Chamorro
>
>[ ] Filipino
>
>[ ] Vietnamese
>
>[ ] Samoan
>
>[ ] Other Asian -- Print race, for example, Hmong, Laotian, Thai, Pakistani, Cambodian, and so on. --> _____________________
>
>[ ] Other Pacific Islander – Print race, for example, Fijian, Tongan, and so on. -->______________________
>
>[ ] Some other race -- Print race. -->
>____________________________________________________

## Ranking with Children

```{r Ranking Single Female with Kids Homeownership2}

ranking(H_earntype_kids,
        'kids',
        plot_title = "Single Earner Female Homeownership with Children",
        #title_scale = 0.8,
        caption_text = 
      "Source: Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```

