---
title: "Gender Equity Dashboard"
output: html_document
    # toc: true
    # toc_float: true
    # code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      dev.args=list(bg="transparent"), fig.width=15, fig.height=12)

library(glptools)
glp_load_packages(TRUE)
library(arrow)
library(viridis)
library(forcats)
library(hrbrthemes)
library(glpdata)


showtext_auto()
font_add("Museo Sans", "MuseoSans_300.otf")

```



```{r, include=FALSE}
#Overlayed histogram
func_plt_hist_overlay <- function(data, var) {

  data$group <- data[[var]]
  
  possible_colors <- c("#0E4A99", "#F58021", "#00A9B7", "#800055", "#356E39", "#CFB94C", "#7E9C80")
  number_I_need = length(unique(data$group))
  these_colors <- possible_colors[1:number_I_need]

  p <- data %>% ggplot( aes(x=HHINCOME, fill=group, color = group, weight = HHWT)) +
    geom_histogram( aes(y=(..count..)/sum(..count..)), 
                   alpha=0.5, position = 'identity', binwidth = 10000 ) +
    scale_fill_manual(values = these_colors) +
    scale_color_manual(values = these_colors) +

    labs(fill="") +
    xlab("Household Income") +
    ylab("Percentage") 
    
  
  return (p)
  
}
  
```

```{r prep_data, eval=FALSE, include=FALSE}
#Skip
# Harrison
#census_data <- read_feather("../../glpdata/data-raw/microdata/acs_micro.feather")

#Josh
#census_data <- read_feather("/Users/joshmarkwell/Desktop/Professional/GLP/projects/gender_equity/raw_data/acs_micro_repwts.feather")


census_data %<>%
  filter(FIPS == "21111")

write_feather(census_data, "raw_data/census_data.feather")


save(population_tract, file = "raw_data/tract_population.RData")

```

# Housing data

In this first chunk, let's read in the Census microdata. Here is some example code on how to read in the data, create new variables to categorize the rows of data into groups, and then summarize the data to create information about Louisville.

Our goal is to create variables for gender, age group, whether someone is a mother, whether someone is married, their level of education, their income, whether they are the head of household, and the number of children they have.



```{r read in new data, eval=FALSE, include=FALSE}
#Skip - only use if changing data to include

#acs_micro_8_11_22 <- read_feather("raw_data/acs_micro_FIPS_repwts_8_11_22.feather") 

census_data %<>%
  filter(FIPS == "21111" | FIPS == '37081')


write_feather(census_data, "raw_data/census_data_8_11_22.feather")

```


```{r explore new data, include=FALSE}
#Start
#File for use below
census_microdata081122 <- read_feather("raw_data/census_data_8_11_22.feather") 

```
This code chunk will identify which households are homeowners vs. renters (in the homeownership variable) and which households are cost-burdened, meaning they pay more thatn 30% of their income toward rent or a mortgage (in the cost_burden variable).

There are also variables for severe cost burden (households that pay more than half of their income towards housing) and households with severe housing problems (lacking a kitchen, adequate plumbing, or an ample number of rooms for the number of people living there).


```{r clean data, include= FALSE}
# To classify the housing data into homeowners, renters, cost-bu
#also create income brackets, earner types

census_microdata081122 %<>%
  group_by(year, SERIAL) %>%
  mutate(hh_members = n()) %>%
  ungroup() %>%
  filter(PERNUM == 1) %>%
  mutate(
    OWNCOST  = replace(OWNCOST, OWNCOST == 99999, NA),
    HHINCOME = replace(HHINCOME, HHINCOME == 99999, NA),
    OWNERSHP = replace(OWNERSHP, OWNERSHP == 0, NA),
    RENTGRS  = replace(RENTGRS, RENTGRS == 0 & OWNERSHP == 1, NA),

    homeownership = if_else(OWNERSHP == 1, T, F),

    hcost = if_else(homeownership == 1, OWNCOST, RENTGRS),
    cost_burden = if_else(hcost * 12 / HHINCOME > 0.3, T, F),
    severe_cost_burden = if_else(hcost * 12 / HHINCOME > 0.5, 1, 0),

    hh_type = case_when(
      homeownership  & !cost_burden ~ "noncb_homeowner",
      homeownership  & cost_burden  ~ "cb_homeowner",
      !homeownership & !cost_burden ~ "noncb_renter",
      !homeownership & cost_burden  ~ "cb_renter",
      TRUE ~ NA_character_),
    
    KITCHEN  = replace(KITCHEN, KITCHEN == 0, NA),
    ROOMS    = replace(ROOMS, ROOMS == 0, NA),
    PLUMBING = replace(PLUMBING, PLUMBING == 0, NA),

    severe_housing_problems = if_else(
      KITCHEN == 1 | PLUMBING == 10 | hh_members / ROOMS > 1 | severe_cost_burden, T, F),
    
#income brackets based on HHINCOME
  #LH-Lower half -> lower40% 
  #UH-Upper Half -> 40th to 80th 
  #T20 -> top 20th percentile

    
    income_group = case_when(
      HHINCOME <= quantile(HHINCOME, probs = 0.4) ~ 'LH',
      HHINCOME <= quantile(HHINCOME, probs = 0.8) ~ 'UH',
      HHINCOME > quantile(HHINCOME, probs = 0.8) ~ 'T20',
      TRUE ~ NA_character_),

  earner_type = case_when(
    HHINCOME == INCTOT ~ 'single_earner',
    HHINCOME != INCTOT ~ 'multi_earner',
    HHINCOME == 0 ~ 'no_earnings'),

  kd_pres = if_else(NCHILD > 0, "kids", "no_kids"), 

    male_fem_mult_earn = case_when(
      sex == 'female' & earner_type == 'single_earner' ~ 'single_fem_earner',
      sex == 'male' & earner_type == 'single_earner' ~ 'single_male_earner',
      earner_type == 'multi_earner' ~ 'multiple_earner') 
)


```

``` {r louisville 2019 data, include = FALSE}

lville_2019 <- census_microdata081122 %>% 
  filter(FIPS == '21111',
         year == '2019')

```


#Waffle Chart
```{r Harrison, include=FALSE}


```


# Homeownership

```{r Trend in Homeownership by Gender}

H_gen_trend<-survey_by_demog(census_microdata081122, weight_var = "HHWT", 'homeownership')

H_gen_trend %<>%
  filter(
    var_type == 'percent',
    race == 'total',
    sex != 'total')

trend(H_gen_trend, 
      homeownership,
      plot_title = "Homeownership by Year", 
      cat = 'sex', 
      y_title = 'Percent',
      caption_text = 
      "Source Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```

```{r Ranking Single Female Homeownership}

H_singFem_rank <- census_microdata081122 %>%
  filter(
    year == '2019',
    earner_type == 'single_earner') %>%
  survey_by_demog('homeownership', weight_var = "HHWT") %>%
  filter(
    sex == 'total',
    race == 'total',
    var_type == 'percent')

ranking(H_singFem_rank, 'homeownership')

```

```{r Ranking Single Female with Kids Homeownership}

H_sinFem_kids <-census_microdata081122 %>%
  filter(
    year == '2019',
    earner_type == 'single_earner',
    NCHILD > 0) %>%
  survey_by_demog('homeownership', weight_var = "HHWT") %>%
  filter(
    sex == 'total',
    race == 'total',
    var_type == 'percent')

ranking(H_sinFem_kids, 'homeownership')

```

```{r Trend in Homeownership by for Single Females with vs. without Kids}

H_s_m_kids_trend <- survey_by_demog(census_microdata081122 ,weight_var = "HHWT", 'homeownership', other_grouping_vars = c("kd_pres"))
  
H_s_m_kids_trend %<>%
  filter(
    var_type == 'percent',
    race == 'total',
    sex == "female") %>%
  pivot_wider(names_from = 'kd_pres', values_from = 'homeownership') %>%
  select(-sex)

trend(H_s_m_kids_trend, 
      kids:no_kids, 
      plot_title = "Female With Kids Homeownership by Year", 
      cat = c("Kids Present" = "kids", "No Kids Present" = "no_kids"), 
      y_title = 'Percent',
      caption_text = 
      "Source Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```

```{r Trend in Female Homeownership by Race}

H_Fem_race <- survey_by_demog(census_microdata081122 ,weight_var = "HHWT", 'homeownership')

H_Fem_race %<>%
  filter(
    var_type == 'percent',
    race != 'total',
    sex == 'total')

trend(H_Fem_race, 
      homeownership, 
      plot_title = "Female With Kids Homeownership by Year", 
      cat = 'race', 
      y_title = 'Percent',
      caption_text = 
      "Source Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```



#Income

```{r, include = FALSE}
#Finding income ranges

#95th perc of income
cut_95 = quantile(lville_2019$HHINCOME, probs = 0.95)

```


```{r 2019 Louisville Men and Women Income Compared}

#################KEEP################

## create hist for each of the most three recent years -> just for verification...that there isnt anything weird
## if feeling funky, use gganimate to create a gif of histograms from 2000 to 2019

#facet this

lville_2019 %>% 
  filter(HHINCOME <= cut_95) %>%
        func_plt_hist_overlay( "sex")
```

```{r 2019 Louisville Single Female Income by Race}

lville_2019 %>% 
  filter(
    sex == 'female',
    earner_type == 'single_earner',
    HHINCOME <= cut_95) %>%
  func_plt_hist_overlay( "race")

```

```{r 2019 Louisville Cost of living Burden for Single Females - stacked histogram}

cost_burden_sf <- lville_2019 %>% 
  filter(
    sex == 'female',
    earner_type == 'single_earner',
    HHINCOME <= cut_95) %>%
  mutate(
    cost_burden = factor(cost_burden, 
                         levels = rev(c(TRUE, FALSE)), 
                         labels = rev(c("Cost Burdened", "Non Cost Burdened")), 
                         ordered = TRUE))

ggplot(cost_burden_sf, 
       aes(x = HHINCOME, fill = cost_burden), 
       color="#00A9B7", 
       alpha=0.5, 
       position = "stack", 
       binwidth = 10000) + 
geom_histogram()

```

```{r Cost-burdened by earner type Trend }

I_CB_earn_trend <- survey_by_demog(census_microdata081122, weight_var = "HHWT", 'cost_burden', other_grouping_vars = c('earner_type'), breakdowns = "sex")

I_CB_earn_trend %<>%
  filter(
    var_type == 'percent',
    race == 'total',
    sex == 'female') %>%
  select(-sex) %>%
  pivot_wider(names_from = "earner_type", values_from = "cost_burden")
  

trend(I_CB_earn_trend, 
      multi_earner:single_earner, 
      plot_title = "Homeownership by Year", 
      cat = c("Multi Earner" = "multi_earner", "Single Earner" = "single_earner"), 
      y_title = 'Percent',
      caption_text = 
      "Source Greater Louisville Project
       Data from GLP analysis of ACS microdata from IPUMS-USA")

```



# Education

```{r Education for Single Male Vs. Single Female}

E_singM_singF <- lville_2019 %>% 
  filter(earner_type == 'single_earner') %>%
  group_by(sex, educ) %>%
  summarize(n=sum(HHWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100,
    educ = factor(educ, 
                  levels = rev(c("no_hs", "hs", "some_col", "assoc", "bach","grad")), 
                  ordered = TRUE))

ggplot(E_singM_singF, 
       aes(x=sex, y=rate, fill=educ)) + 
  geom_bar(stat="identity", position='fill')

```

```{r, Education of Single Female by race}

E_singF_race <- lville_2019 %>% 
  filter(
    sex == 'female',
    earner_type == 'single_earner') %>%
  group_by(race, educ) %>%
  summarize(n=sum(HHWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100,
    educ = factor(educ, 
                  levels = rev(c("no_hs", "hs", "some_col", "assoc", "bach","grad")), 
                  ordered = TRUE))

ggplot(E_singF_race, aes(x=race, y=rate, fill=educ)) + 
geom_bar(stat="identity", position='fill')

```


# Household Age
```{r Cost Burdened Status by Age}

#Stacked histogram

cost_burden_age_sf <- census_microdata081122 %>% 
    filter(year %in% 2010:2019) %>%
  mutate(
    cost_burden = factor(cost_burden, 
                         levels = rev(c(TRUE, FALSE)), 
                         labels = rev(c("Cost Burdened", "Non Cost Burdened")), 
                         ordered = TRUE),
    
    age_group = case_when(
      age %in% 15:19 ~ NA_character_, age %in% 20:29 ~ "20 - 29", 
      age %in%  30:39 ~ "30 - 39",  age %in% 40:49 ~ "40 - 49",  age %in% 50:59 ~ "50 - 59",  age %in% 60:69 ~ "60 - 69", 
      age %in% 70:79 ~ "70 - 79", age >= 80 ~ "80 + "
    )
    
    # male_fem_mult_earn = case_when(
    #   sex == 'female' & earner_type == 'single_earner' ~ 'single_fem_earner',
    #   sex == 'male' & earner_type == 'single_earner' ~ 'single_male_earner',
    #   earner_type == 'multi_earner' ~ 'multiple_earner')
    )

ggplot(cost_burden_age_sf,
       aes(x=age_group, y=HHWT , fill=cost_burden),
        color="#00A9B7") +
geom_bar(stat="identity", position='fill')


##Remove NA values from cost_burden_age_sf dataframe
```

```{r  Cost Burdened Stage by Age/Race}

ggplot(cost_burden_age_sf,
       aes(x=age_group, y=HHWT , fill=cost_burden),
        color="#00A9B7") +
geom_bar(stat="identity", position='fill')+
facet_wrap(~male_fem_mult_earn)

```

# Earner Types Over Time
```{r Proportion of Earner Types over time}

earner_trend <- census_microdata081122 %>%
  
  mutate(
    male_fem_mult_earn = case_when(
      sex == 'female' & earner_type == 'single_earner' ~ 'single_fem_earner',
      sex == 'male' & earner_type == 'single_earner' ~ 'single_male_earner',
      earner_type == 'multi_earner' ~ 'multiple_earner')
  ) %>% 
  group_by(year, male_fem_mult_earn) %>%
  summarize(n=sum(HHWT, na.rm = TRUE)) %>%
  mutate(
    total = sum(n),
    rate = n/sum(n)*100) 
  
ggplot(earner_trend, 
       aes(x=year, y=rate, fill=male_fem_mult_earn),
        color="#00A9B7") + 
geom_bar(stat="identity", position='fill')
```


# Mortgage Data
